Declarative Deployment Reconfiguration of Pulverised Collective Adaptive Systems at Runtime \thanks{ This work has been supported by the Italian PRIN Project COMMON-WEARS (2020HCWWLP) and the EU/MUR FSE REACT-EU PON R\&I 2014-2020. }

### 2023-05-21 paper-2023-acsos-pulverization.tex
RW: needs to mention how our approach fits or differs @nicolasfara @danysk some bits - pls check

In our approach, the application partitioning into components is manually specified through a \ac{dsl}.
%
Therefore, it can be framed in the context of \emph{component-based software engineering},
 with ~\cite{vale2016component-based-se} providing a comprehensive review.
%
The specification or design of distributed systems
 may leverage
 component models~\cite{DBLP:journals/tse/CrnkovicSVC11},
 architectural description languages~\cite{DBLP:journals/tse/MedvidovicT00},
 service-based compositions~\cite{DBLP:journals/csur/LemosDB16}, 
 or frameworks.
%
Specifically,
 this work builds on the \emph{pulverisation} component model~\cite{FI2020-pulverization,IEEE-IoTJ-pulverization-simulation}.
 whereby a large-scale cyber-physical system
 is partitioned into a graph of devices
 where each device is split into five deployable components:
 (i) sensors interface,
 (ii) actuators interface,
 (iii) behaviour,
 (iv) state, and
 (v) communication component.
%
In~\cite{IEEE-IoTJ-pulverization-simulation},
 a methodology on top of the pulverisation model
 is proposed
 for generating and testing deployments through simulation.

Additionally,
 the approach of specifying in a single codebase 
 parts of the structure, behaviour, interaction 
 and/or other aspects of a whole distributed system
 is also related to \emph{macroprogramming}~\cite{Casadei2023macro}
 and, in particular, to \emph{multi-tier programming} paradigm~\cite{DBLP:journals/csur/WeisenburgerWS20},
 where the deployment units for different tiers
(e.g., client and server tiers; or view, business logic, and data tiers)
are obtained by compilation or interpretation of a single codebase.

>>>>

In our approach, the application partitioning into components is manually specified through a \ac{dsl}.
%
Therefore, it can be framed in the context of \emph{component-based software engineering},
 with ~\cite{vale2016component-based-se} providing a comprehensive review.
%
The specification or design of distributed systems
 may leverage
 component models~\cite{DBLP:journals/tse/CrnkovicSVC11},
 architectural description languages~\cite{DBLP:journals/tse/MedvidovicT00},
 service-based compositions~\cite{DBLP:journals/csur/LemosDB16}, 
 or frameworks.
%
Specifically,
 this work builds on the \emph{pulverisation} component model~\cite{FI2020-pulverization,IEEE-IoTJ-pulverization-simulation}.
 whereby a large-scale cyber-physical system
 is partitioned into a graph of devices
 where each device is split into five deployable components:
 (i) sensors interface,
 (ii) actuators interface,
 (iii) behaviour,
 (iv) state, and
 (v) communication component.
%
In~\cite{IEEE-IoTJ-pulverization-simulation},
 a methodology on top of the pulverisation model
 is proposed
 for generating and testing deployments through simulation.
%
There, file descriptors and scripts are used to generate deployment plans; in \ourframework{}, instead, we provide \acp{dsl} supporting a specification-oriented, declarative approach.
%

Additionally,
 the approach of specifying in a single codebase 
 parts of the structure, behaviour, interaction 
 and/or other aspects of a whole distributed system
 is also related to \emph{macroprogramming}~\cite{Casadei2023macro}
 and, in particular, to \emph{multi-tier programming} paradigm~\cite{DBLP:journals/csur/WeisenburgerWS20},
 where the deployment units for different tiers
(e.g., client and server tiers; or view, business logic, and data tiers)
are obtained by compilation or interpretation of a single codebase.
%
\ourframework{} adopts the same idea,
 where a logically centralised specification
 of the distributed architecture
 enables runtime checks and deployment activities.
